---
title: 'CosMx Discovery: Manual Cluster Annotation'
author: "CosMx-Nf"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
    df_print: paged
  pdf_document:
    toc: true
    toc_depth: 3
    latex_engine: xelatex
    geometry: landscape, margin=1.5cm
params:
  run_id: default
  cluster_basis: leiden_clus_res0.3
  giotto_dir: ''
  top_n: 10
  python_path: ''
  overwrite_labels: false
  error_on_incomplete: false
  compare_cols: "insitutype_unsup,insitutype_sup_KPMP,insitutype_sup_HCA_Kidney,ht_call"
  subcluster: "3"
  subcluster_res: 0.5
  b_lineage: true
  b_min_genes: 2
  b_min_counts: 80
  b_res: 0.6
  if_control: ""
---

Explores the **discovery** output of run `run_id` (the merged Giotto object, before annotation)
and writes the `cluster_id,label` CSV that the `annotate` stage (`bin/apply_labels.R`) consumes.

Paths follow `run.sh`: every run lives beside this repository in `Runs/<run_id>/results/`. Re-knit for another run or resolution by overriding params, e.g.
`rmarkdown::render("manual_annotation.Rmd", params = list(run_id = "full", cluster_basis = "leiden_clus_res0.8"))`.

```{r setup, include=FALSE}

# PDF has no code folding, so echoing there buries the figures in source listings
TO_PDF <- identical(knitr::opts_knit$get("rmarkdown.pandoc.to"), "latex")

knitr::opts_chunk$set(echo = !TO_PDF, message = FALSE, warning = FALSE,
                      dev = "png", dpi = 200, fig.width = 8, fig.height = 6,
                      out.width = "100%", fig.align = "center")

```

```{r Environment}

library(Giotto)
library(Matrix)
library(tidyverse)

Sys.setenv(KMP_DUPLICATE_LIB_OK = "TRUE")
if (nzchar(params$python_path)) Sys.setenv(RETICULATE_PYTHON = params$python_path)

find_pipe <- function(d = getwd()) {
  while (!file.exists(file.path(d, "main.nf"))) {
    if (dirname(d) == d) {
      stop("main.nf not found above ", getwd(), " -- open this notebook from inside the Pipeline repo")
    }
    d <- dirname(d)
  }
  normalizePath(d)
}

PIPE_DIR    <- find_pipe()
RES_DIR     <- file.path(dirname(PIPE_DIR), "Runs", params$run_id, "results")
# Subcluster columns live on the subcluster object, Leiden columns on the discover one
IS_SUB      <- startsWith(params$cluster_basis, "sub_")
GIOTTO_DIR  <- if (nzchar(params$giotto_dir)) params$giotto_dir else
  file.path(RES_DIR, "objects", if (IS_SUB) "subcluster" else "discover", "giotto")
DIAG_DIR    <- file.path(RES_DIR, "tables", if (IS_SUB) "05_subclustering" else "03_clustering", "markers")
LABELS_OUT  <- file.path(RES_DIR, "tables", "06_annotation", "labels",
                         str_c("cluster_labels.", params$cluster_basis, ".csv"))
MARKERS_CSV <- file.path(PIPE_DIR, "assets", "canonical_markers.csv")
CLUSTER_COL <- params$cluster_basis

if (!dir.exists(GIOTTO_DIR)) stop("No discover object for run '", params$run_id, "' at ", GIOTTO_DIR)

# S13b re-clusters a parent cluster, so it has nothing to do when the basis IS the subclusters
SUB_OK <- nzchar(params$subcluster) && !startsWith(CLUSTER_COL, "sub_")

```

```{r Helpers}

theme_pixel <- function() theme_bw(base_size = 11) + theme(panel.grid = element_blank())

scale_diverging <- function(...) {
  scale_fill_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0, ...)
}

read_diag <- function(suffix) {
  f <- file.path(DIAG_DIR, str_c(CLUSTER_COL, suffix))
  if (file.exists(f)) read_csv(f, show_col_types = FALSE) else NULL
}

# Mean expression and fraction of cells detecting each gene, as genes x clusters matrices
cluster_stats <- function(m, groups) {
  ind <- t(fac2sparse(groups))
  n   <- colSums(ind)
  list(mean = sweep(as.matrix(m %*% ind), 2, n, "/"),
       pct  = sweep(as.matrix((m > 0) %*% ind), 2, n, "/"))
}

z_by_gene <- function(mat) {
  z <- t(scale(t(mat)))
  z[!is.finite(z)] <- 0
  z
}

to_long <- function(mat, value) {
  mat |>
    as_tibble(rownames = "gene") |>
    pivot_longer(-gene, names_to = "cluster", values_to = value)
}

fig_dim  <- function(n, per, lo, hi = 40) max(lo, min(hi, n * per))
fig_rows <- function(n, n_facets = 1, per = 0.15, pad = 1.5, hi = 60) {
  min(hi, n * per + n_facets * 0.15 + pad)
}

```

# S0 · Load discovery object

```{r Load}

g <- loadGiotto(GIOTTO_DIR)

cells <- getCellMetadata(g, output = "data.table") |> as_tibble()
norm  <- getExpression(g, values = "normalized", output = "matrix")
stopifnot(CLUSTER_COL %in% names(cells))

# Leiden ids sort numerically (2 before 10), not as text
clusters <- str_sort(unique(as.character(cells[[CLUSTER_COL]])), numeric = TRUE)
cells <- cells |>
  mutate(cluster = factor(as.character(.data[[CLUSTER_COL]]), levels = clusters))
cell_cluster <- cells$cluster[match(colnames(norm), cells$cell_ID)]

markers <- read_csv(MARKERS_CSV, show_col_types = FALSE) |>
  filter(gene %in% rownames(norm))
markers_by_ct   <- split(markers$gene, markers$cell_type)
markers_by_comp <- split(markers$gene, markers$compartment)
canon_genes     <- unique(markers$gene)

cat("run:", params$run_id, "\nlabels:", LABELS_OUT, "\n")
cat(str_glue("Cells: {nrow(cells)}   Genes: {nrow(norm)}   {CLUSTER_COL}: {length(clusters)} clusters"), "\n")

```

```{r Figure sizes}

n_clu   <- length(clusters)
w_clu   <- fig_dim(n_clu, 0.45, 9, 20)
h_ctpan <- fig_rows(max(lengths(markers_by_ct)), per = 0.30, pad = 1.3)
h_gene  <- fig_rows(length(canon_genes), length(markers_by_comp))
h_ct    <- fig_rows(length(markers_by_ct), per = 0.30)
h_clu   <- fig_dim(n_clu, 0.35, 6)
h_vol   <- fig_dim(ceiling(n_clu / 3), 2.6, 6)

```

# S1 · Overview & QC recap

```{r Samples}

count(cells, sample_id)

```

```{r Treatment by timepoint}

if (all(c("treatment", "timepoint") %in% names(cells))) {
  cells |>
    count(treatment, timepoint) |>
    pivot_wider(names_from = timepoint, values_from = n)
}

```

```{r QC recap, fig.height=4}

qc_cols <- intersect(c("nCount_RNA", "nFeature_RNA", "total_expr", "nr_feats"), names(cells))

if (length(qc_cols)) {
  cells |>
    select(sample_id, all_of(qc_cols)) |>
    pivot_longer(-sample_id) |>
    ggplot(aes(sample_id, value, fill = sample_id)) +
    geom_violin(scale = "width") +
    facet_wrap(~name, scales = "free_y") +
    theme_pixel() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") +
    labs(x = NULL, y = NULL, title = "Per-cell QC by sample")
}

```

# S2 · Clustering overview across resolutions (Viz 1)

```{r Cluster summary}

cluster_summary <- file.path(RES_DIR, "tables", "03_clustering", "cluster_summary.csv")
if (file.exists(cluster_summary)) read_csv(cluster_summary, show_col_types = FALSE)

```

```{r UMAP ladder, fig.width=7, fig.height=6, results='asis'}

ladder <- str_sort(str_subset(names(cells), "^leiden_clus_res"), numeric = TRUE)

walk(ladder, \(col) {
  cat("\n\n###", col, "\n\n")
  plotUMAP(g, dim_reduction_name = "umap", cell_color = col,
           point_size = 0.8, show_plot = FALSE, return_plot = TRUE) |>
    print()
})

```

# S3 · Working resolution

```{r Working resolution}

count(cells, cluster)

```

# S4 · Per-cluster marker discovery (Viz 2)

Reuses the pipeline's pre-computed scran markers when available, otherwise recomputes with the
exact `bin/diagnostics.R` call.

```{r Markers}

de  <- read_diag("_de_all.csv")
top <- read_diag("_de_top.csv")

if (is.null(de)) {
  message("No pre-computed diagnostics for ", CLUSTER_COL, " -- recomputing findMarkers_one_vs_all()")
  de <- findMarkers_one_vs_all(g, method = "scran", expression_values = "normalized",
                               cluster_column = CLUSTER_COL, min_feats = 10) |>
    as_tibble()
}
de <- de |> mutate(cluster = factor(as.character(cluster), levels = clusters))

if (is.null(top)) top <- de |> arrange(ranking) |> slice_head(n = params$top_n, by = cluster)

es_col  <- str_subset(names(de), regex("logFC|LFC|logfoldchange|difference|summary", ignore_case = TRUE))[1]
sig_col <- intersect(c("FDR", "padj", "p.value", "p_val", "pval"), names(de))[1]
cat("effect-size column:", coalesce(es_col, "<none>"), " | significance column:", coalesce(sig_col, "<none>"), "\n")

sel_genes <- de |>
  filter(if (is.na(sig_col)) TRUE else .data[[sig_col]] < 0.05) |>
  arrange(ranking) |>
  slice_head(n = params$top_n, by = cluster) |>
  pull(feats) |>
  unique() |>
  intersect(rownames(norm))
h_hm <- fig_rows(length(sel_genes))

head(top, 20)

```

# S5 · Effect-size heatmap of top markers (Viz 3)

```{r Top marker heatmap, fig.width=w_clu, fig.height=h_hm}

top_z <- cluster_stats(norm[sel_genes, , drop = FALSE], cell_cluster)$mean |> z_by_gene()
gene_order <- rownames(top_z)[order(max.col(top_z, "first"))]

top_z |>
  to_long("z") |>
  mutate(gene = factor(gene, levels = gene_order), cluster = factor(cluster, levels = clusters)) |>
  ggplot(aes(cluster, gene, fill = z)) +
  geom_tile() +
  scale_diverging() +
  theme_pixel() +
  theme(axis.text.y = element_text(size = 8), axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(fill = "z(mean)", title = str_c("Top ", params$top_n, " markers per cluster"))

```

# S6 · Per-cluster volcano plots (Viz 4)

```{r Volcano, fig.width=10, fig.height=h_vol}

if (!is.na(es_col) && !is.na(sig_col)) {
  volcano <- de |>
    filter(is.finite(.data[[es_col]]), is.finite(.data[[sig_col]])) |>
    mutate(x   = .data[[es_col]],
           y   = -log10(pmax(.data[[sig_col]], 1e-300)),
           hit = .data[[sig_col]] < 0.05 & abs(x) > 1)
  labelled <- volcano |> arrange(ranking) |> slice_head(n = params$top_n, by = cluster)

  gene_labels <- if (requireNamespace("ggrepel", quietly = TRUE)) {
    ggrepel::geom_text_repel(data = labelled, aes(label = feats), size = 2, max.overlaps = 20)
  } else {
    geom_text(data = labelled, aes(label = feats), size = 2, check_overlap = TRUE)
  }

  ggplot(volcano, aes(x, y)) +
    geom_point(aes(colour = hit), size = 0.4) +
    geom_vline(xintercept = c(-1, 1), linetype = 2, linewidth = 0.3) +
    gene_labels +
    scale_colour_manual(values = c(`FALSE` = "grey70", `TRUE` = "#B2182B"), guide = "none") +
    facet_wrap(~cluster, scales = "free", ncol = 3) +
    theme_pixel() +
    labs(x = es_col, y = "-log10(sig)")
} else {
  cat("DE table lacks effect-size and/or p-value columns -- showing top markers by rank instead.\n")
  top |> select(cluster, feats, ranking) |> arrange(cluster, ranking) |> knitr::kable()
}

```

# S7 · Canonical markers, one heatmap per cell type (Viz 5)

One heatmap per cell type, grouped by compartment. **Fill = mean expression z-scored once per
gene across all clusters** (so colour is comparable between panels; clipped at the 99th
percentile), **number = % of cells expressing** (blank below 10%). A gene that marks several
cell types appears under each. Use S13 to find which type a cluster resembles, then this to see
which genes drove it.

```{r Canonical marker heatmaps, fig.width=w_clu, fig.height=h_ctpan, results='asis'}

canon <- cluster_stats(norm[canon_genes, , drop = FALSE], cell_cluster)

canon_stats <- to_long(canon$mean, "mean_expr") |>
  left_join(to_long(canon$pct, "pct"), by = join_by(gene, cluster)) |>
  mutate(z = as.numeric(scale(mean_expr)), .by = gene) |>
  mutate(z = if_else(is.finite(z), z, 0), cluster = factor(cluster, levels = clusters))
lim <- quantile(abs(canon_stats$z), 0.99, names = FALSE)

panels <- markers |>
  distinct(compartment, cell_type, gene) |>
  arrange(compartment, cell_type) |>
  nest(genes = gene, .by = c(compartment, cell_type)) |>
  mutate(new_compartment = compartment != lag(compartment, default = ""))

pwalk(panels, \(compartment, cell_type, genes, new_compartment) {
  if (new_compartment) cat("\n\n##", compartment, "\n")
  cat("\n\n###", cell_type, "\n\n")
  p <- canon_stats |>
    filter(gene %in% genes$gene) |>
    mutate(gene = factor(gene, levels = rev(genes$gene))) |>
    ggplot(aes(cluster, gene, fill = z)) +
    geom_tile(colour = "white", linewidth = 0.4) +
    geom_text(aes(label = if_else(pct >= 0.10, as.character(round(pct * 100)), "")), size = 2.1) +
    scale_diverging(limits = c(-lim, lim), oob = scales::squish) +
    theme_pixel() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1), axis.text.y = element_text(size = 8)) +
    labs(x = CLUSTER_COL, y = NULL, fill = "z(mean)",
         title = str_glue("{cell_type}  ({compartment}, {nrow(genes)} markers)"))
  print(p)
})

```

# S8 · Feature plots by compartment (Viz 6)

```{r Feature plots, fig.width=10, fig.height=8, results='asis'}

iwalk(markers_by_comp, \(genes, compartment) {
  cat("\n\n### ", compartment, "\n\n")
  dimFeatPlot2D(g, feats = head(unique(genes), 12), dim_reduction_to_use = "umap",
                dim_reduction_name = "umap", expression_values = "normalized",
                cow_n_col = 4, point_size = 0.4, show_plot = FALSE, return_plot = TRUE) |>
    print()
})

```

# S9 · Violin plots per cluster (Viz 7)

```{r Violins, fig.width=10, fig.height=6, results='asis'}

iwalk(markers_by_comp, \(genes, compartment) {
  cat("\n\n### ", compartment, "\n\n")
  violinPlot(g, feats = head(unique(genes), 8), cluster_column = CLUSTER_COL,
             expression_values = "normalized", show_plot = FALSE, return_plot = TRUE) |>
    print()
})

```

# S10 · Mean-marker heatmap (Viz 8)

```{r Mean marker heatmap, fig.width=w_clu, fig.height=h_gene}

plotMetaDataHeatmap(g, expression_values = "normalized", metadata_cols = CLUSTER_COL,
                    selected_feats = canon_genes, show_plot = FALSE, return_plot = TRUE)

```

# S11 · Cluster composition (Viz 9)

```{r Composition by sample, fig.height=5}

cells |>
  count(sample_id, cluster) |>
  mutate(prop = n / sum(n), .by = sample_id) |>
  ggplot(aes(sample_id, prop, fill = cluster)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  theme_pixel() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = NULL, y = "fraction of cells", title = "Cluster composition per sample")

```

```{r Composition by treatment, fig.height=5}

if ("treatment" %in% names(cells)) {
  cells |>
    count(treatment, cluster) |>
    mutate(prop = n / sum(n), .by = treatment) |>
    ggplot(aes(treatment, prop, fill = cluster)) +
    geom_col() +
    scale_y_continuous(labels = scales::percent) +
    theme_pixel() +
    labs(x = NULL, y = "fraction of cells", title = "Cluster composition per treatment")
}

```

# S12 · Leiden vs InSituType cross-check (Viz 11)

```{r Leiden vs InSituType, fig.width=8, fig.height=h_clu}

if ("insitutype_unsup" %in% names(cells)) {
  cells |>
    mutate(insitutype_unsup = as.character(insitutype_unsup),
           insitutype_unsup = factor(insitutype_unsup, levels = str_sort(unique(insitutype_unsup), numeric = TRUE))) |>
    count(cluster, insitutype_unsup) |>
    mutate(prop = n / sum(n), .by = cluster) |>
    ggplot(aes(insitutype_unsup, cluster, fill = prop)) +
    geom_tile() +
    geom_text(aes(label = round(prop, 2)), size = 2) +
    scale_fill_viridis_c() +
    theme_pixel() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(x = "insitutype_unsup", y = CLUSTER_COL, title = "Row-normalized cluster overlap")
} else {
  cat("insitutype_unsup not present in metadata.\n")
}

```

# S13 · Marker-score-guided labeling

```{r Marker scores, fig.width=w_clu, fig.height=h_ct}

# Same scoring as bin/diagnostics.R, so these match the pipeline's template CSV
MIN_DETECTION   <- 0.05  # Rarer genes have a tiny sd, so z turns noise into specificity
MIN_PANEL_GENES <- 3     # A 2-gene panel can still win a cluster on one loud gene
MIN_SCORE       <- 0.5   # Below this the winner is only the least-negative, so no call

detected  <- rowMeans(norm[canon_genes, , drop = FALSE] > 0)
genes_use <- canon_genes[detected >= MIN_DETECTION]

gene_z <- canon$mean[genes_use, , drop = FALSE] |> z_by_gene() |> to_long("z")

scores <- markers |>
  distinct(cell_type, gene) |>
  inner_join(gene_z, by = join_by(gene), relationship = "many-to-many") |>
  summarise(score = round(mean(z), 3), .by = c(cell_type, cluster)) |>
  mutate(cluster = factor(cluster, levels = clusters))

n_used <- markers |>
  filter(gene %in% genes_use) |>
  distinct(cell_type, gene) |>
  count(cell_type, name = "n_genes_used")
eligible <- n_used |> filter(n_genes_used >= MIN_PANEL_GENES) |> pull(cell_type)
cat("cell types eligible for best_guess:", length(eligible), "of", length(markers_by_ct), "\n")

ggplot(scores, aes(cluster, cell_type, fill = score)) +
  geom_tile() +
  scale_fill_viridis_c() +
  theme_pixel() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), axis.text.y = element_text(size = 9)) +
  labs(title = "Canonical-marker enrichment per cluster (z-scored per gene)", fill = "z score")

```

```{r Label template}

# Margin is the gap to the runner-up. Small margin + high best_score: two sibling types tie.
# Small margin + best_score near 0: nothing is enriched, so best_guess is left blank.
top_markers <- top |>
  arrange(ranking) |>
  summarise(top_markers = str_c(head(feats, 5), collapse = ";"), .by = cluster) |>
  mutate(cluster_id = as.character(cluster), .keep = "unused")

template <- scores |>
  filter(cell_type %in% eligible) |>
  arrange(cluster, desc(score)) |>
  summarise(best_guess   = cell_type[1],
            second_guess = cell_type[2],
            best_score   = score[1],
            margin       = round(score[1] - score[2], 3),
            .by = cluster) |>
  mutate(best_guess = if_else(best_score < MIN_SCORE, "", best_guess)) |>
  left_join(n_used, by = join_by(best_guess == cell_type)) |>
  left_join(count(cells, cluster, name = "n_cells"), by = join_by(cluster)) |>
  mutate(cluster_id = as.character(cluster), .keep = "unused") |>
  left_join(top_markers, by = join_by(cluster_id)) |>
  mutate(label = "") |>
  select(cluster_id, n_cells, best_guess, second_guess, best_score, margin,
         n_genes_used, top_markers, label)

no_call <- template |> filter(best_guess == "") |> pull(cluster_id)
if (length(no_call)) cat("no panel match (best_guess blank) for cluster(s):", str_flatten_comma(no_call), "\n")

suggested <- template |> select(cluster_id, best_guess) |> deframe()

arrange(template, margin)

```

# S13b · Subcluster

Re-clusters the cells of `subcluster` on their own PCA (all genes, no batch correction), so a
sample that separates inside the cluster shows up as its own subcluster.
Set `subcluster` to a cluster id (comma-separated for several) to run this section.

```{r Subcluster, eval=SUB_OK}

SUB_DIMS <- 1:30
SUB_K    <- 20

sub_ids   <- str_split_1(params$subcluster, ",") |> str_trim()
sub_cells <- cells |>
  filter(cluster %in% sub_ids) |>
  pull(cell_ID)

# An empty subset reaches runPCA as a request for <= 0 components, which errors obscurely
if (!length(sub_cells))
  stop("subcluster '", params$subcluster, "' matched no cells. Available: ",
       str_flatten_comma(levels(cells$cluster)))

g_sub <- subsetGiotto(g, cell_ids = sub_cells)
g_sub <- runPCA(g_sub, feat_type = "rna", feats_to_use = NULL, ncp = max(SUB_DIMS), name = "pca_sub")
g_sub <- createNearestNetwork(g_sub, feat_type = "rna", type = "sNN", dim_reduction_to_use = "pca",
                              dim_reduction_name = "pca_sub", dimensions_to_use = SUB_DIMS, k = SUB_K,
                              name = "NN.pca_sub")
g_sub <- doLeidenCluster(g_sub, feat_type = "rna", nn_network_to_use = "sNN", network_name = "NN.pca_sub",
                         resolution = params$subcluster_res, name = "subcluster", seed_number = 42)
g_sub <- runUMAP(g_sub, feat_type = "rna", dim_reduction_to_use = "pca", dim_reduction_name = "pca_sub",
                 dimensions_to_use = SUB_DIMS, name = "umap_sub", seed_number = 42)

sub_meta <- getCellMetadata(g_sub, output = "data.table") |> as_tibble()
cat(length(sub_cells), "cells ->", n_distinct(sub_meta$subcluster), "subclusters\n")

```

```{r Subcluster by sample, eval=SUB_OK}

# Percent of each subcluster coming from each sample
sub_meta |>
  count(subcluster, sample_id) |>
  mutate(pct = round(100 * n / sum(n)), .by = subcluster) |>
  select(-n) |>
  pivot_wider(names_from = sample_id, values_from = pct, values_fill = 0) |>
  left_join(count(sub_meta, subcluster, name = "cells"), by = join_by(subcluster)) |>
  arrange(subcluster)

```

```{r Subcluster UMAP, fig.width=8, fig.height=6, eval=SUB_OK}

plotUMAP(g_sub, dim_reduction_name = "umap_sub", cell_color = "sample_id",
         point_size = 0.3, show_plot = FALSE, return_plot = TRUE)

plotUMAP(g_sub, dim_reduction_name = "umap_sub", cell_color = "subcluster", show_center_label = TRUE,
         point_size = 0.3, show_plot = FALSE, return_plot = TRUE)

```

```{r Subcluster markers, eval=SUB_OK}

findMarkers_one_vs_all(g_sub, method = "scran", expression_values = "normalized",
                       cluster_column = "subcluster", min_feats = 10) |>
  as_tibble() |>
  arrange(ranking) |>
  slice_head(n = 8, by = cluster) |>
  summarise(top_markers = str_c(feats, collapse = ", "), .by = cluster)

```

Immune panels from `canonical_markers.csv` on the subcluster UMAP. Score = fraction of the
panel's genes detected in each cell; `Immune` is every gene in the `immune_infiltrate` compartment.

```{r Subcluster immune panels, fig.width=8, fig.height=6, eval=SUB_OK}

IMMUNE_PANELS <- c("B_cell", "T_cell", "Treg")

sub_norm <- getExpression(g_sub, values = "normalized", output = "matrix")

immune_panels <- c(markers_by_ct[IMMUNE_PANELS],
                   Immune = list(markers$gene[markers$compartment == "immune_infiltrate"])) |>
  map(\(genes) intersect(unique(genes), rownames(sub_norm)))

sub_umap <- getDimReduction(g_sub, reduction = "cells", reduction_method = "umap",
                            name = "umap_sub", output = "matrix")

sub_scores <- immune_panels |>
  map(\(genes) colMeans(sub_norm[genes, , drop = FALSE] > 0)) |>
  as_tibble() |>
  mutate(cell_ID = colnames(sub_norm)) |>
  right_join(tibble(cell_ID = rownames(sub_umap), UMAP1 = sub_umap[, 1], UMAP2 = sub_umap[, 2]),
             by = join_by(cell_ID))

walk(names(immune_panels), \(panel) {
  p <- sub_scores |>
    arrange(.data[[panel]]) |>
    ggplot(aes(UMAP1, UMAP2, colour = .data[[panel]])) +
    geom_point(size = 0.3) +
    scale_colour_viridis_c() +
    theme_pixel() +
    labs(title = str_glue("{panel}: fraction of {length(immune_panels[[panel]])} markers detected"),
         colour = NULL)
  print(p)
})

```

```{r Subcluster immune genes, fig.width=12, fig.height=12, eval=SUB_OK}

dimFeatPlot2D(g_sub, feats = unique(unlist(immune_panels[IMMUNE_PANELS])),
              expression_values = "normalized", dim_reduction_to_use = "umap",
              dim_reduction_name = "umap_sub", order = TRUE, cow_n_col = 4, point_size = 0.3,
              show_plot = FALSE, return_plot = TRUE)

```

# S13c · B/plasma lineage subcluster

Follows Danaher et al., *Sci Transl Med* 2024 (PMID 39602512): pull the B and plasma cells out,
subcluster them on their own, and filter to `b_min_counts` transcripts for the embedding.
They classified with a CD20 immunofluorescence channel we do not have, so membership here is a
multi-gene rule instead -- MS4A1 alone tops out at 3 transcripts in this dataset.

```{r B lineage cells, eval=params$b_lineage}

B_GENES <- c("MS4A1", "CD79A", "CD79B", "CD19", "BANK1", "PAX5", "CR2", "CD22",
             "TCL1A", "VPREB3", "MZB1", "JCHAIN", "XBP1", "IRF4", "TNFRSF17",
             "CD27", "CD38", "IGHM", "IGHD", "IGKC", "IGHG1/2")

raw_mat  <- getExpression(g, values = "raw", output = "matrix")
b_use    <- intersect(B_GENES, rownames(raw_mat))
b_hits   <- colSums(raw_mat[b_use, , drop = FALSE] > 0)
n_counts <- colSums(raw_mat)

b_cells <- cells |>
  mutate(b_genes = b_hits[cell_ID], total_counts = n_counts[cell_ID]) |>
  filter(b_genes >= params$b_min_genes, total_counts > params$b_min_counts)

cat(length(b_use), "of", length(B_GENES), "B-lineage genes on the panel:",
    str_flatten_comma(b_use), "\n")
cat(nrow(b_cells), "cells with >=", params$b_min_genes, "B genes and >",
    params$b_min_counts, "counts\n\n")

count(b_cells, cluster, name = "cells")

```

```{r B lineage subcluster, eval=params$b_lineage}

B_DIMS <- 1:20

g_b <- subsetGiotto(g, cell_ids = b_cells$cell_ID)
g_b <- runPCA(g_b, feat_type = "rna", feats_to_use = NULL, ncp = max(B_DIMS), name = "pca_b")
g_b <- createNearestNetwork(g_b, feat_type = "rna", type = "sNN", dim_reduction_to_use = "pca",
                            dim_reduction_name = "pca_b", dimensions_to_use = B_DIMS, k = 15,
                            name = "NN.pca_b")
g_b <- doLeidenCluster(g_b, feat_type = "rna", nn_network_to_use = "sNN", network_name = "NN.pca_b",
                       resolution = params$b_res, name = "b_sub", seed_number = 42)
g_b <- runUMAP(g_b, feat_type = "rna", dim_reduction_to_use = "pca", dim_reduction_name = "pca_b",
               dimensions_to_use = B_DIMS, name = "umap_b", seed_number = 42)

b_meta <- getCellMetadata(g_b, output = "data.table") |> as_tibble()
cat(nrow(b_meta), "cells ->", n_distinct(b_meta$b_sub), "subclusters\n")

```

Composition by treatment and timepoint. Percentages are of each sample's own cells, which is what makes them
comparable -- the samples differ several-fold in cell count and in depth.

```{r B lineage composition, eval=params$b_lineage}

b_meta |>
  count(sample_id, treatment, timepoint, name = "b_cells") |>
  left_join(count(cells, sample_id, name = "total"), by = join_by(sample_id)) |>
  mutate(pct_of_sample = round(100 * b_cells / total, 3)) |>
  arrange(treatment, timepoint)

```

```{r B lineage by subcluster, eval=params$b_lineage}

b_meta |>
  count(b_sub, sample_id) |>
  mutate(pct = round(100 * n / sum(n)), .by = b_sub) |>
  select(-n) |>
  pivot_wider(names_from = sample_id, values_from = pct, values_fill = 0) |>
  left_join(count(b_meta, b_sub, name = "cells"), by = join_by(b_sub)) |>
  arrange(b_sub)

```

IF intensity per subcluster. These are the channels InSituType already cohorts on; CD45 is the
leukocyte readout that does not depend on the RNA panel at all.

```{r B lineage IF, eval=params$b_lineage}

if_cols <- str_subset(names(b_meta), "^(Mean|Max)\\.")

if (length(if_cols)) {
  b_meta |>
    summarise(across(all_of(if_cols), \(v) round(median(v, na.rm = TRUE), 1)),
              cells = n(), .by = b_sub) |>
    arrange(b_sub)
} else {
  cat("no IF/morphology columns on the object -- check names(pDataDT(g))\n")
}

```

```{r B lineage UMAP, fig.width=8, fig.height=6, eval=params$b_lineage}

plotUMAP(g_b, dim_reduction_name = "umap_b", cell_color = "b_sub", show_center_label = TRUE,
         point_size = 1.2, show_plot = FALSE, return_plot = TRUE)

plotUMAP(g_b, dim_reduction_name = "umap_b", cell_color = "treatment",
         point_size = 1.2, show_plot = FALSE, return_plot = TRUE)

```

```{r B lineage markers, eval=params$b_lineage}

findMarkers_one_vs_all(g_b, method = "scran", expression_values = "normalized",
                       cluster_column = "b_sub", min_feats = 5) |>
  as_tibble() |>
  arrange(ranking) |>
  slice_head(n = 10, by = cluster) |>
  summarise(top_markers = str_c(feats, collapse = ", "), .by = cluster)

```

```{r B lineage genes, fig.width=12, fig.height=10, eval=params$b_lineage}

dimFeatPlot2D(g_b, feats = b_use, expression_values = "normalized",
              dim_reduction_to_use = "umap", dim_reduction_name = "umap_b", order = TRUE,
              cow_n_col = 4, point_size = 0.8, show_plot = FALSE, return_plot = TRUE)

```

# S14 · Draft annotated UMAP (Viz 10)

```{r Draft UMAP, fig.width=8, fig.height=6}

g <- annotateGiotto(g, annotation_vector = suggested, cluster_column = CLUSTER_COL,
                    name = "cell_type_draft")

plotUMAP(g, dim_reduction_name = "umap", cell_color = "cell_type_draft",
         show_center_label = TRUE, point_size = 0.3, show_plot = FALSE, return_plot = TRUE)

```

# S15 · Write the `cluster_id,label` CSV

`label_map` is seeded from `best_guess` so the notebook knits to a draft; **replace it with your
curated labels after reviewing S5-S14.** Copy the scaffold below to get an editable literal.
Once the CSV exists it is kept as-is: set `overwrite_labels: true` to write your edits.

```{r Label scaffold, class.source='fold-show'}

cat("label_map <- c(\n",
    str_flatten(str_glue('    "{names(suggested)}" = "{suggested}"'), ",\n"),
    "\n)\n", sep = "")

```

```{r Label map, class.source='fold-show'}

# EDIT: replace the seed below with your reviewed labels (paste the scaffold above), e.g.
# label_map <- c("1" = "PT", "2" = "CNT_CD", "3.1" = "Fibroblast", "3.2" = "Macrophage")
label_map <- suggested

```

```{r Write labels}

if (file.exists(LABELS_OUT) && !params$overwrite_labels) {
  message("labels file exists; keeping it (set overwrite_labels = TRUE to replace). Reading it back.")
  labels <- read_csv(LABELS_OUT, col_types = cols(.default = col_character()))
} else {
  labels <- enframe(label_map, name = "cluster_id", value = "label")
  dir.create(dirname(LABELS_OUT), recursive = TRUE, showWarnings = FALSE)
  write_csv(labels, LABELS_OUT, na = "")
  message("Wrote ", LABELS_OUT)
}

labels

```

# S16 · Validate against the annotate contract

Mirrors `bin/apply_labels.R` -- every distinct value of `cluster_basis` must have a non-blank label.

```{r Validate, class.source='fold-show'}

labelled_ids <- labels |> filter(!is.na(label), label != "") |> pull(cluster_id)
unlabelled   <- setdiff(clusters, labelled_ids)

if (length(unlabelled)) {
  cat("[!] UNLABELED clusters:", str_flatten_comma(unlabelled), "\n")
  if (params$error_on_incomplete) stop("Incomplete labels for ", CLUSTER_COL)
} else {
  cat("[OK] All", length(clusters), "clusters in", CLUSTER_COL, "are labeled.\n")
}

cat("\nLaunch the annotate stage with:\n")
cat(str_glue("  cd {PIPE_DIR} && ./run.sh {params$run_id} --stage annotate ",
             "--labels {LABELS_OUT} --cluster_basis {CLUSTER_COL} -profile server"), "\n")

```

# S17 · Review figures

Every gene in `REVIEW_PANELS` is plotted, unlike S7-S9 which cap each panel. Dot size is the
percentage of cells detecting the gene, fill is mean expression z-scored per gene across
clusters. Genes absent from the CosMx panel are listed rather than silently dropped.

```{r Review panels}

REVIEW_PANELS <- list(
  Plasma      = c("MZB1", "JCHAIN", "IGKC", "IGHG1/2", "XBP1", "TNFRSF17", "IRF4"),
  B_cell      = c("MS4A1", "CD79A", "CD79B", "CD19", "BANK1", "CD22", "TCL1A", "VPREB3"),
  T_NK        = c("CD3D", "CD3E", "CD2", "IL7R", "TRAC", "CD8A", "NKG7", "GNLY", "KLRD1"),
  Myeloid     = c("CD68", "CD14", "C1QA", "C1QB", "LYZ", "TYROBP", "FCER1G", "ITGAM"),
  Leukocyte   = c("PTPRC", "CD52", "LAPTM5", "CORO1A", "CXCR4"),
  PT          = c("LRP2", "CUBN", "ALDOB", "GPX3", "GATM", "SLC34A1", "ASS1"),
  PT_injured  = c("VCAM1", "HAVCR1", "CLU", "SPP1", "CRYAB"),
  TAL_DCT     = c("UMOD", "SLC12A1", "SLC12A3", "EGF", "KNG1"),
  CNT_CD      = c("CALB1", "AQP2", "AQP3", "SCNN1G", "SLC4A1", "SLC26A4", "DEFB1"),
  Podocyte    = c("PODXL", "NPHS1", "NPHS2", "PLA2R1", "VEGFA"),
  Endothelial = c("PECAM1", "EMCN", "VWF", "FLT1", "KLF2", "EPAS1"),
  Stromal     = c("DCN", "LUM", "COL1A1", "C7", "MGP", "SELENOP"),
  Mural       = c("ACTA2", "MYH11", "TAGLN", "RGS5", "PDGFRB", "REN")
)

review <- tibble(panel = names(REVIEW_PANELS), gene = REVIEW_PANELS) |>
  unnest(gene) |>
  mutate(panel = fct_inorder(panel))

absent <- review |> filter(!gene %in% rownames(norm))
if (nrow(absent)) {
  cat("not on the CosMx panel:", str_flatten_comma(absent$gene), "\n")
}
review <- review |> filter(gene %in% rownames(norm))
h_rev  <- max(6, 0.22 * nrow(review) + 1.5)

cat(nrow(review), "genes across", n_distinct(review$panel), "panels\n")

```

```{r Review dotplot, fig.width=w_clu, fig.height=h_rev}

rev_stats <- cluster_stats(norm[review$gene, , drop = FALSE], cell_cluster)

to_long(z_by_gene(rev_stats$mean), "z") |>
  left_join(to_long(rev_stats$pct, "pct"), by = join_by(gene, cluster)) |>
  left_join(review, by = join_by(gene), relationship = "many-to-many") |>
  mutate(gene = fct_rev(fct_inorder(gene)), cluster = factor(cluster, levels = clusters)) |>
  ggplot(aes(cluster, gene)) +
  geom_point(aes(size = pct, fill = z), shape = 21, colour = "grey40", stroke = 0.2) +
  scale_size_area(max_size = 5, labels = scales::percent) +
  scale_diverging() +
  facet_grid(panel ~ ., scales = "free_y", space = "free_y", switch = "y") +
  theme_pixel() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        axis.text.y = element_text(size = 7),
        strip.text.y.left = element_text(angle = 0, size = 8),
        strip.placement = "outside") +
  labs(x = CLUSTER_COL, y = NULL, size = "% detected", fill = "z(mean)")

```

## Depth per cluster

A cluster whose cells carry fewer transcripts cannot be typed as confidently as the rest, so
the depth distribution has to be shown before any claim about what does or does not resolve.

```{r Depth per cluster, fig.width=w_clu, fig.height=5}

depth_col <- intersect(c("total_expr", "nCount_RNA"), names(cells))[1]
med_all   <- median(cells[[depth_col]])

cells |>
  mutate(lab = median(.data[[depth_col]]), .by = cluster) |>
  ggplot(aes(cluster, .data[[depth_col]])) +
  geom_violin(aes(fill = lab < med_all), scale = "width", linewidth = 0.2) +
  geom_boxplot(width = 0.14, outlier.shape = NA, fill = "white", linewidth = 0.3) +
  geom_hline(yintercept = med_all, linetype = 2, colour = "grey40") +
  scale_y_log10(labels = scales::comma) +
  scale_fill_manual(values = c(`FALSE` = "#9EC9BE", `TRUE` = "#B2182B"), guide = "none") +
  theme_pixel() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x = CLUSTER_COL, y = "transcripts per cell",
       title = "Transcripts per cell by cluster",
       subtitle = str_glue("red = median below the dataset median ({med_all}); dashed = that median"))

```

```{r Depth table}

cells |>
  summarise(n = n(),
            median = median(.data[[depth_col]]),
            q25 = quantile(.data[[depth_col]], 0.25),
            q75 = quantile(.data[[depth_col]], 0.75),
            .by = cluster) |>
  arrange(median)

```

## Is there a B-cell compartment?

MS4A1 is CD20. If its positive cells were B cells they would also carry CD79A/CD19/BANK1 well
above the background rate. Detection tracks depth, so the comparison is made within deciles of
total counts -- a naive fold would report depth as biology.

```{r B cell evidence}

B_OTHER <- c("CD79A", "CD19", "BANK1", "CD79B", "CD22")

raw_mat <- getExpression(g, values = "raw", output = "matrix")
b_other <- intersect(B_OTHER, rownames(raw_mat))

ev <- tibble(cell_ID = colnames(raw_mat),
             counts  = colSums(raw_mat),
             genes   = colSums(raw_mat > 0),
             ms4a1   = as.numeric(raw_mat["MS4A1", ]),
             other_b = colSums(raw_mat[b_other, , drop = FALSE] > 0)) |>
  mutate(decile = ntile(counts, 10))

cat("MS4A1 raw counts across all cells:\n")
print(table(ev$ms4a1))

ev |>
  mutate(group = case_when(ms4a1 >= 2 ~ "MS4A1 >= 2",
                           ms4a1 == 1 ~ "MS4A1 = 1",
                           .default   = "MS4A1 = 0")) |>
  summarise(cells = n(), median_counts = median(counts), median_genes = median(genes),
            pct_other_B = round(100 * mean(other_b > 0), 1), .by = group) |>
  arrange(desc(cells))

```

MS4A1 against every other gene on the panel. Detection rate alone is not the argument: the
comparison that matters is **peak count in a single cell**, because a real population gives a
loud signal in few cells while ambient contamination gives a whisper spread across many.

```{r Gene ranking}

# rowMax without densifying: a 6K panel over ~150k cells is ~7 GB dense
row_max <- function(m) {
  s <- as(m, "TsparseMatrix")
  v <- numeric(nrow(m))
  o <- tapply(s@x, s@i + 1L, max)
  v[as.integer(names(o))] <- o
  v
}

gene_stats <- tibble(gene = rownames(raw_mat),
                     pct_cells = 100 * rowSums(raw_mat > 0) / ncol(raw_mat),
                     mean_per_cell = rowSums(raw_mat) / ncol(raw_mat),
                     max_in_a_cell = row_max(raw_mat)) |>
  mutate(rank_detection = rank(-pct_cells))

REFS <- c("MS4A1", "CD79A", "CD79B", "CD19", "BANK1", "CD22",
          "MZB1", "JCHAIN", "IGKC", "TNFRSF17",
          "PTPRC", "CD68", "CD3E", "ALDOB", "PODXL", "ACTA2", "B2M")

gene_stats |>
  filter(gene %in% REFS) |>
  arrange(desc(pct_cells)) |>
  mutate(across(c(pct_cells, mean_per_cell), \(v) round(v, 3)),
         rank_detection = str_c(rank_detection, " of ", nrow(gene_stats)))

```

```{r Gene ranking summary}

gene_stats |>
  summarise(across(c(pct_cells, max_in_a_cell),
                   list(median = \(v) round(median(v), 2),
                        p90    = \(v) round(quantile(v, 0.9), 2),
                        p10    = \(v) round(quantile(v, 0.1), 2))))

```

```{r B cell depth matched}

# Expected co-detection = the MS4A1-negative rate in the same depth decile
ev |>
  summarise(pos = sum(ms4a1 > 0),
            obs = sum(ms4a1 > 0 & other_b > 0),
            bg  = mean(other_b[ms4a1 == 0] > 0), .by = decile) |>
  summarise(observed = sum(obs), expected = round(sum(pos * bg), 1)) |>
  mutate(fold = round(observed / expected, 2))

```

```{r B cell depth plot, fig.width=9, fig.height=4}

ev |>
  mutate(group = if_else(ms4a1 > 0, "MS4A1+", "MS4A1-")) |>
  ggplot(aes(group, counts, fill = group)) +
  geom_violin(scale = "width") +
  geom_boxplot(width = 0.12, outlier.shape = NA, fill = "white") +
  scale_y_log10(labels = scales::comma) +
  theme_pixel() +
  theme(legend.position = "none") +
  labs(x = NULL, y = "total counts per cell",
       title = "MS4A1 is detected in deep cells, not in B cells")

```

# S18 · Immunofluorescence check

CD45 intensity is measured from the image, so it is independent of transcript depth -- the one
readout in this dataset that the depth confound cannot reach. `if_control` names a cluster that
must be CD45-negative; its 95th percentile sets the null, so the null column reads ~5%.
Left blank it defaults to the largest cluster.

```{r IF channels}

IF_COLS <- str_subset(names(cells), "^(Mean|Max)\\.")

if (!length(IF_COLS)) {
  cat("no IF columns on this object -- check names(pDataDT(g))\n")
} else {
  cat(length(IF_COLS), "channels:", str_flatten_comma(IF_COLS), "\n")
}

CTRL <- if (nzchar(params$if_control)) params$if_control else
  count(cells, cluster) |> slice_max(n, n = 1) |> pull(cluster) |> as.character()
cat("CD45-negative control cluster:", CTRL, "\n")

```

```{r IF by cluster, eval=length(IF_COLS) > 0}

null_95 <- cells |> filter(cluster == CTRL) |> pull(Mean.CD45) |> quantile(0.95, na.rm = TRUE)

cells |>
  summarise(n = n(),
            CD45  = round(median(Mean.CD45, na.rm = TRUE)),
            CD68  = round(median(Mean.CD68, na.rm = TRUE)),
            PanCK = round(median(Mean.PanCK, na.rm = TRUE)),
            pct_CD45_above_null = round(100 * mean(Mean.CD45 > null_95, na.rm = TRUE), 1),
            .by = cluster) |>
  arrange(desc(pct_CD45_above_null))

```

Intensity is not comparable between slides. When treatment is confounded with slide, any
treatment-level claim has to be made within a slide.

```{r IF by slide, eval=length(IF_COLS) > 0}

cells |>
  summarise(CD45 = round(median(Mean.CD45, na.rm = TRUE)), n = n(),
            .by = c(cluster, slide_id)) |>
  pivot_wider(names_from = slide_id, values_from = c(CD45, n), values_fill = 0) |>
  arrange(cluster)

```

```{r IF plot, fig.width=w_clu, fig.height=6, eval=length(IF_COLS) > 0}

ord <- cells |>
  summarise(m = median(Mean.CD45, na.rm = TRUE), .by = cluster) |>
  arrange(desc(m)) |>
  pull(cluster)

cells |>
  mutate(cluster = factor(as.character(cluster), levels = as.character(ord))) |>
  ggplot(aes(cluster, Mean.CD45)) +
  geom_violin(aes(fill = cluster), scale = "width", linewidth = 0.2) +
  geom_boxplot(width = 0.12, outlier.shape = NA, fill = "white", linewidth = 0.3) +
  geom_hline(yintercept = null_95, linetype = 2, colour = "#B2182B") +
  scale_y_log10(labels = scales::comma) +
  theme_pixel() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none") +
  labs(x = CLUSTER_COL, y = "Mean CD45 intensity",
       title = str_glue("CD45 protein per cluster (dashed = 95th pct of cluster {CTRL})"))

```

# S16c · Compare the annotations

Puts the manual labels, the unsupervised InSituType clusters and each supervised pass side by
side. Columns in `compare_cols` that are not on the object are skipped.

```{r Comparison setup}

compare_cols <- str_trim(str_split_1(params$compare_cols, ","))
skipped      <- setdiff(compare_cols, names(cells))
compare_cols <- intersect(compare_cols, names(cells))
if (length(skipped)) cat("not on the object, skipped:", str_flatten_comma(skipped), "\n")

# Manual labels are only in the CSV until --stage annotate runs
if (file.exists(LABELS_OUT)) {
  manual <- read_csv(LABELS_OUT, col_types = cols(.default = col_character())) |> deframe()
  cells <- cells |> mutate(manual = unname(manual[as.character(cluster)]))
  compare_cols <- c("manual", compare_cols)
} else {
  cat("no labels CSV at", LABELS_OUT, "-- manual column skipped\n")
}
cat("comparing:", str_flatten_comma(compare_cols), "\n")

```

## Composition

```{r Comparison composition, results='asis'}

walk(compare_cols, \(k) {
  d <- cells |>
    filter(!is.na(.data[[k]])) |>
    count(across(all_of(k)), sort = TRUE) |>
    mutate(pct = round(100 * n / sum(n), 1))
  cat(str_glue("\n\n**{k}** -- {nrow(d)} labels\n\n"))
  print(knitr::kable(head(d, 25)))
})

```

## Agreement

```{r Comparison agreement}

# Adjusted Rand index: 0 = chance, 1 = identical partitions
ari <- function(a, b) {
  ok <- !is.na(a) & !is.na(b)
  t  <- table(a[ok], b[ok])
  s  <- function(x) sum(choose(x, 2))
  e  <- s(rowSums(t)) * s(colSums(t)) / choose(sum(t), 2)
  m  <- (s(rowSums(t)) + s(colSums(t))) / 2
  round((s(t) - e) / (m - e), 3)
}

pairs <- combn(compare_cols, 2, simplify = FALSE)

pairs |>
  map(\(p) tibble(a = p[1], b = p[2],
                  ARI = ari(cells[[p[1]]], cells[[p[2]]]),
                  n_shared = sum(!is.na(cells[[p[1]]]) & !is.na(cells[[p[2]]])))) |>
  list_rbind() |>
  knitr::kable()

```

## Cross-tabs

```{r Comparison cross-tabs, fig.width=11, fig.height=8}

walk(pairs, \(p) {
  p_tab <- cells |>
    filter(!is.na(.data[[p[1]]]), !is.na(.data[[p[2]]])) |>
    count(a = .data[[p[1]]], b = .data[[p[2]]]) |>
    mutate(frac = n / sum(n), .by = a)
  p_plot <- ggplot(p_tab, aes(b, a, fill = frac)) +
    geom_tile() +
    scale_fill_gradient(low = "white", high = "#B2182B", limits = c(0, 1)) +
    theme_pixel() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(title = str_c(p[1], " (rows) vs ", p[2], " (cols)"), x = p[2], y = p[1], fill = "row frac")
  print(p_plot)
})

```

## Merge table

```{r Merge table}

# One row per observed label combination, largest first; fill merged_label to build a labels CSV
merge_table <- cells |>
  count(across(all_of(compare_cols)), sort = TRUE) |>
  mutate(pct = round(100 * n / sum(n), 2), merged_label = "")

merge_out <- file.path(dirname(LABELS_OUT), "annotation_comparison.csv")
write_csv(merge_table, merge_out, na = "")
cat("wrote", merge_out, "--", nrow(merge_table), "combinations\n\n")

knitr::kable(head(merge_table, 40))

```
